---
title: "Lab Getting Started"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

``` {r, echo=FALSE}
	library(readr)
	library("ggplot2")

cpsa2016_cpsaat03_clean <- read_delim("GettingStarted/cpsa2016_cpsaat03_clean.csv", 
	                                      "\t", escape_double = FALSE, trim_ws = TRUE)
	#View(cpsa2016_cpsaat03_clean)
```

# Summary Statistics

``` {r, echo=FALSE}
	summary(cpsa2016_cpsaat03_clean)
```

# Scatter Plot 1

Scatter Plot of Age and Civilian Labor Force by Percent of Population

``` {r, echo=FALSE}
	
	ggplot(data=cpsa2016_cpsaat03_clean, aes(x=Age, y=CLF.Pct)) +
	  geom_point() + 
	  theme(axis.text.x=element_text(angle=90,hjust=1)) +
	  ggtitle("Age vs. Civilian Labor Force by Percent of Population") +
	  labs(x = "Age", y ="Civilian Labor Force Percent") 
```

# Bar Chart 1

Bar Chart of Civilian Labor Force Total with labels by Age

``` {r, echo=FALSE}
	ggplot(data=cpsa2016_cpsaat03_clean, aes(x=Age, y=CLF.Total)) + 
	  geom_bar(stat="identity") + 
	  theme(axis.text.x=element_text(angle=90,hjust=1)) +
	  ggtitle("Civilian Labor Force Total by Age") +
	  labs(x = "Age", y ="Civilian Labor Force Total")
```

# Scatter Plot 2

Scatter Plot of Civilian Labor Force (Percent), Age, and Civilian Labor Force (Employed) 

``` {r, echo=FALSE}
	ggplot(cpsa2016_cpsaat03_clean, aes(x = Age, y = CLF.Total, col = CLF.Pct)) +
	  geom_point(size=2) + scale_color_gradient(low="lightpink",high = "darkred") +
	  theme(axis.text.x=element_text(angle=90,hjust=1)) +
	  ggtitle("Age vs. Civilian Labor Force Total") +
	  labs(x = "Age", y ="Civilian Labor Force Total")
```

# Scatter Plot 3

Scatter Plot of Percent of Population Not in Labor Force by Age with green dots for Ages with labor force   non-participation over 50%

``` {r, echo=FALSE}	

	cpsa2016_cpsaat03_clean$Colorvariable <-cut(cpsa2016_cpsaat03_clean$NILF.Pct, breaks=c(-Inf, 50, Inf), labels=c("black", "green"))
	ggplot(data=cpsa2016_cpsaat03_clean) +
	  geom_point(mapping=aes(x=Age, y=NILF.Pct), color=cpsa2016_cpsaat03_clean$Colorvariable) + 
	  theme(axis.text.x=element_text(angle=90,hjust=1)) +
	  ggtitle("Age vs. Percent of Population Not in Labor Force") +
	  labs(x = "Age", y ="Percent of Population Not in Labor Force")
```